# Solutions {.unnumbered}

## Assignment 1: Two-Sample Tests & Permutation Tests

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(ISLR2)
data("College")
```

### Question 1: Room and Board Costs

**(a) Boxplot and histograms**

```{r}
ggplot(College, aes(y = Room.Board, x = Private, fill = Private)) +
  geom_boxplot() +
  ggtitle("Room and Board by Private/Public")
```

```{r}
ggplot(College, aes(x = Room.Board, fill = Private)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
  ggtitle("Room and Board by Private/Public")
```

**(b) Is the t-test justified?**

The histograms show approximately normal distributions for both groups, with similar shapes. The two-sample t-test is justified here.

**(c) Two-sample t-test**

```{r}
College %>%
  group_by(Private) %>%
  summarize(
    mean = mean(Room.Board),
    sd = sd(Room.Board),
    n = n()
  )
```

```{r}
t.test(Room.Board ~ Private, data = College)
```

The variances differ between groups, so we use the Welch t-test (R's default). The p-value is very small, indicating a significant difference in room and board costs between private and public colleges.

**(d) One-sided test**

```{r}
t.test(Room.Board ~ Private, data = College, alternative = "greater")
```

### Question 2: Top 10 Percent & Permutation Test

**(a) Boxplots and histograms**

```{r}
ggplot(College, aes(y = Top10perc, x = Private, fill = Private)) +
  geom_boxplot() +
  ggtitle("Top10perc by Private/Public")
```

```{r}
ggplot(College, aes(x = Top10perc, fill = Private)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
  ggtitle("Top10perc by Private/Public")
```

**(b) Observed t-statistic**

```{r}
College %>%
  group_by(Private) %>%
  summarize(
    mean = mean(Top10perc),
    sd = sd(Top10perc)
  )
```

```{r}
observed_t <- t.test(Top10perc ~ Private, data = College)$statistic
observed_t
```

**(c) Normality assessment**

The histograms show right-skewed distributions, particularly for private colleges. The normality assumption is not well-satisfied. A permutation test is a better alternative since it makes no distributional assumptions.

**(d) Permutation test**

```{r}
set.seed(191)
n_iterations <- 10000
t_stat_shuffle <- numeric(n_iterations)

for (i in 1:n_iterations) {
  shuffled_private <- sample(College$Private)
  t_stat_shuffle[i] <- t.test(Top10perc ~ shuffled_private, data = College)$statistic
}
```

```{r}
ggplot(data.frame(t = t_stat_shuffle), aes(x = t)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = observed_t, color = "red", linewidth = 1) +
  ggtitle("Permutation Null Distribution vs. Observed t-statistic") +
  labs(x = "t-statistic", y = "Count")
```

```{r}
perm_pvalue <- mean(abs(t_stat_shuffle) >= abs(observed_t))
cat("Permutation p-value:", perm_pvalue, "\n")
```

The permutation p-value confirms the parametric result: there is a significant difference in `Top10perc` between private and public colleges, and this conclusion holds without assuming normality.

### Question 3: Sign Test

**(a)**

```{r}
data("Fund")
manager14_returns <- Fund$Manager14
n_positive <- sum(manager14_returns > 0)
cat("Number of positive months for Manager14:", n_positive, "out of 50\n")
```

**(b)**

```{r}

```

**(c)**

```{r}

```

**(d)**

```{r}

```

**(e)**

```{r}

```

**(f)**

```{r}

```

### Question 4: One-Way ANOVA

**(a)**

```{r}
data("Wage")
```

**(b)**

```{r}

```

**(c)**

```{r}

```

**(d)**

```{r}

```

**(e)**

```{r}

```

**(f)**

```{r}

```
